Reasoning about a Capability Machine with Local Capabilities

### 2017-06-16 popl-2018/paper.tex
removed comment on limitations of local capabilities as it may not be correct.
First attempt at small argument/local capability discussion

\textbf{A word on local capabilities} Local capabilities are an essential part
of the calling convention. The use of local capabilities is, however, not
limitted to the stack and return pointer.

@@

\emph{Modularity} It is important that our calling convention is modular, i.e.
we do not assume that our code is specially privileged w.r.t. the adversary, and
they can apply the same measures to protect themselves from us as we do to
protect ourselves from them. More concretely, the requirements we have on
callbacks and return pointers received from the adversary are also satisfied by
callbacks and return pointers that we pass to them. For example, our return
pointers are local capabilities because they must point to memory where we can
store the old stack pointer, but the adversary's return pointers are also
allowed to be local. Adversary callbacks on the other hand are required to
be global but the callbacks we construct ourselves are allocated on the heap and
also global. \\\\

>>>>

@@

\emph{Modularity} It is important that our calling convention is modular, i.e.
we do not assume that our code is specially privileged w.r.t. the adversary, and
they can apply the same measures to protect themselves from us as we do to
protect ourselves from them. More concretely, the requirements we have on
callbacks and return pointers received from the adversary are also satisfied by
callbacks and return pointers that we pass to them. For example, our return
pointers are local capabilities because they must point to memory where we can
store the old stack pointer, but the adversary's return pointers are also
allowed to be local. Adversary callbacks on the other hand are required to
be global but the callbacks we construct ourselves are allocated on the heap and
also global.

\emph{Arguments and local capabilities} 
%Stack and return pointer
%Local capabilities cannot be "freely used"
%Safety dictated by LR
%Need to dynamical check whether we overwrite our local stuff.
Local capabilities are a central part of the calling convention as they are used
to construct stack and return pointer. The use of local capabilities for the
calling convention unfortunately limits the extend of which local capabilties
can be used for other things. Say we are using the calling convention and
receive a local capability other than the stack and return pointer, then we need
to be careful if we want to use it because it may be an alias to the stack
pointer. That is, if we first push something to the stack and then write to the
local capability, then we may overwrite our local state. While this sounds
unsafe, the safety actually depends on what our memory invariants we want to
establish, so the safety is actually dictated by the logical relation. This may,
however, not be a satisfactory requirement for say compiler developers as it is
impossible to figure out what local capabilities are safe to use. A different
approach would be to do dynamic checks on local capabilities, but this may be to
costly to do in practice.

%Pass capability on
%Stack example
%Pass argument by reference
%Sofisticated deep safety analysis
We also need to be careful when we receive a capability from an adversary that
we want to pass on to a different adversary. It turns out that the logical
relation again dictates when this is safe. Namely, the logical relation says
that we can only pass on safe arguments. For instance, when we receive a stack
pointer from an adversary, then we may at some point want to pass on part of
this stack pointer to say a callback. In order to do so, we need to make sure
the stack pointer is safe which means that it must only point to safe
values. When received from an adversary, we have to consider the contents of the
stack unsafe, so before we pass it on we clear which corresponds to filling it
with safe values as integers are always safe to pass on. In general, if we want
to pass on a local capability we received from an adversary, we have to make
sure it is safe, for instance by clearing everything it points to (if possible,
the capability might be a readonly capability). Clearing everthing is, however,
not always a desirable approach as we may have an array or data structure with
important values we want to pass on by reference in which case it is
unacceptable to just clear the entire thing. In this case, it might be possible
to make a deep safety analysis to make sure that the capability points to
nothing dangerous. What we are left with is a too conservative approach or an
non-obvious expensive dynamic safety analysis.

% Can local capabilities be used for other things?/Local capabilities only for stack?
%Not unheard of - other machine also have things specific for language features.
The use of local capabilities for other things than stack and return pointer
turns out to be somewhat limitted. It is, however, not unusual that machines
have builtin special constructs that are used for just one language feature. A
machine could for instance have call and return instructions with the sole
prupose of handling everything related to calls.
\\\\
% Old version of the above:
% Carefully constructing stack and return
% pointer plays a big part in the calling convention, but it is completely
% undermined if an unsafe argument, say the full stack pointer that include the
% caller's local stack frame, is passed as an argument of the call. The logical
% relation dictates when it is safe to pass an argument to an adversary, namely it
% must be a safe value. In order to illustrate what this means for local
% capabilities, consider the stack capability. When the stack capability is
% received from an adversary, then the contents of the stack is unknown and must
% thus be considered unsafe, so before the stack capability can be passed on, we
% must make sure that it only gives access to safe values. In order to meet this
% requirement, we take a conservative approcah and simply clear everything on the
% stack. Integers are always safe, so the stack capability can now safely be
% passed on. While this approach works for the stack, it may in practice be a bit
% too naive for other local capabilities received form an adversary. Say we
% receive a capability from an adversary that happens to be the first part of the
% stack. We want to pass this capability on, so we naively clear the contents it
% gives access to which in the worst case means we overwrite our local data or the
% activation record. While this is strictly speaking not unsafe (integers are
% always) safe, it is not desirable in a real system. We would in principle also
% like to be able to pass on arguments by reference, e.g a local capability for
% some array of values. In this case, it would not be satisfactory to just clear the
% entire thing. However, the capability still needs to be safe, so there would
% need to be some kind of deep dynamic check that ensures that everything the
% capability gives access to is indeed safe. In a real setting, this may again be
% undesirable.
% This also points to a more
% general problem: When can we use local capabilities we have received from an
% adversary. We at least need to do some kind of dynamic range check to make sure
% that we do not inadverdently overwrite our local stack frame.
